### 2017-11-30 Body/Abstract.tex
FEUP $(date)

\abstract
The behaviour of concrete-filled steel tubes (CFSTs) is governed by the contribution of both the steel tube and the concrete core, and their interaction may lead to the development of multi-axial stress states, namely the concrete confinement effect and the steel bi-axial stress state. From the previous conducted experimental and numerical study on the behaviour of circular CFST members, it was found that the combination of multi-axial stress states will have a positive effects on the bending capacity of CFST columns. And the multi-axial stress states will cause the underestimation of the analytical results of Distributed Plasticity model. To account for the material strength under multi-axial stress state, strength correction methods, which was developed based on 3D Finite Element (FE) model, had been proposed. This was achieved with a comprehensive parametric study, which considered multiple combinations of steel tube diameter-to-thickness ratios, concrete-to-steel strengths and axial load levels. A section factor ? was used to represent both the geometry and material properties of the CFST members, and ?-dependent correction equations were derived. The feasibility of the equations was verified to be good for both the design and simplified modelling of the CFST members.