### 2017-11-30 Body/Abstract.tex
FEUP $(date)

\abstract
The behaviour of concrete-filled steel tubes (CFSTs) is governed by the contribution of both the steel tube and the concrete core. Their interaction may lead to the development of multi-axial stress states, namely the concrete confinement effect and the steel bi-axial stress state. From the previous conducted experimental and numerical studies on the flexural behaviour of circular CFST members, it was found that the combination of multi-axial stress states will benefit the bending capacity of CFST column. Thus, the multi-axial stress states will cause the underestimation on the bending capacity of circular CFST members from the analytical results of Distributed Plasticity (DP) model. To account for the material strength under multi-axial stress state and improve the accuracy of the DP model, strength correction methods was developed based on 3D Finite Element (FE) model. And with a parametric study, the material strength correction equations to account for the multi-axial stress states had been derived. A section factor ? was used to represent both the geometry and material properties of the CFST members, and ?-dependent correction equations were derived. The feasibility of the equations was verified to be good for both the design and simplified modelling of the CFST members.